Study 2 Analysis

### 2024-10-04 study2.qmd
update plots

---
title: "Study 2 Analysis"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
---

@@

```{r}
#| fig.width: 11
#| fig.height: 9

@@

outlier_plot <- ggplot(outliers, aes(x = state, y = z_score, color = is_outlier)) +
  geom_jitter(alpha = 0.5) +
  facet_wrap(~ pct_goal) +
  labs(
    title = "Z-Scores of Change in Energy Usage by Appliance",
    x = "Appliance",
    y = "Z-Score",
    color = "Outlier"
  ) +
  theme_minimal()

outlier_plot

sum1 |> ggplot(aes(x=pct_change)) +
  geom_density(aes(fill=refClass), alpha=0.5) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.1, alpha=0.3) +
  facet_wrap(~state, scales="free_y") +
  theme_minimal() +
  labs(title="Distribution of Percent Change by State and Reference Class",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2")

@@

# Shapiro-Wilk test for normality (sample size limitation)
shapiro_test <- shapiro.test(sum1$pct_change[1:5000])  # Using first 5000 observations if large dataset
print(shapiro_test)

sum1 |> 
  filter(!id %in% outlier_id) |>
  ggplot(aes(x=pct_change)) +
  geom_density(aes(fill=refClass), alpha=0.5) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.1, alpha=0.3) +
  facet_wrap(~state, scales="free_y") +
  theme_minimal() +
  labs(title="Distribution of Percent Change by State and Reference Class",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2")

@@

@@

```{r}
# plot distribution of reductions by applicance
s2_long |> ggplot(aes(x = change)) + geom_histogram(bins = 30) + facet_wrap(~appliance, scales = "free")

DT::datatable(sum1)

DT::datatable(sum1 |> group_by(id,block) |> summarise(diff(total_kWh),diff(orig_kWh)))

sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
  geom_boxplot(position = "dodge") +
  facet_wrap(~state)

 sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") 
```

@@

sum1 |> 
  ungroup() |> 
  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
  mutate(id=reorder(id,pct_change)) |> 
  ggplot(aes(y=id,x=pct_change)) + 
  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
  ggh4x::facet_nested_wrap(refClass~pct_goal,axes="all",scales="free",ncol=2)

@@

@@

# mixed effects model testing effect of refClass, state, and pct_goal on pct_change

m1 <- lmer(pct_change ~ refClass  + (1+plan|id) + (1|state), data = sum1)
summary(m1)

@@

>>>>

---
title: "Study 2 Analysis"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format: 
  html:
    page-layout: full
---

@@

## Inspect response distribution for outliers

@@

outlier_plot <- ggplot(outliers, aes(x = state, y = z_score, color = is_outlier)) +
  geom_jitter(alpha = 0.5) +
  facet_wrap(~ pct_goal) +
  labs(
    title = "Z-Scores of Change in Energy Usage by State - outliers colored",
    x = "State",
    y = "Z-Score",
    color = "Outlier"
  ) +
  theme_minimal()

@@

# Shapiro-Wilk test for normality (sample size limitation)
shapiro_test <- shapiro.test(sum1$pct_change[1:5000])  # Using first 5000 observations if large dataset
print(shapiro_test)
```

::: {.column-page-inset-right}
```{r}
#| layout-ncol: 1
#| fig.width: 12
#| fig.height: 11

#outlier_plot

sum1 |> ggplot(aes(x=pct_change)) +
  geom_density(aes(fill=state), alpha=0.5) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.2, alpha=0.3) +
  facet_wrap(~refClass) +
  theme_minimal() +
  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2")

@@

sum1 |> 
  filter(!id %in% outlier_id) |>
  ggplot(aes(x=pct_change)) +
  geom_density(aes(fill=state), alpha=0.5) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.2, alpha=0.3) +
  facet_wrap(~refClass) +
  theme_minimal() +
  labs(title="After outlier exclusion",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2")

@@

## Individual Participant responses

@@

:::

## Interactive data table

```{r}
# plot distribution of reductions by applicance
# s2_long |> ggplot(aes(x = change)) + geom_histogram(bins = 30) + facet_wrap(~appliance, scales = "free")
DT::datatable(sum1)
#DT::datatable(sum1 |> group_by(id,block) |> summarise(diff(total_kWh),diff(orig_kWh)))
```

## Main Effects and Interactions

```{r}
#| layout-ncol: 2

sum1 |> ggplot(aes(x = refClass, y = pct_change, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")

sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = pct_goal)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")

sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")

 sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")

sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") + facet_wrap(~state)

sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = rounded)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") + facet_wrap(~rounded)

@@

# mixed effects model testing effect of refClass, state, and pct_goal on pct_change
# marginal effects plots
library(ggeffects)

m1 <- lmer(pct_change ~ refClass  + (1+plan|id) + (1|state), data = sum1)
summary(m1)
ggpredict(m1, c("refClass")) |> plot()

m1 <- lmer(pct_change ~ pct_goal  + (1|id) + (1|state), data = sum1)
summary(m1)
ggpredict(m1, c("pct_goal")) |> plot()

m1 <- lmer(pct_change ~ pct_goal*rounded  + (1|id) + (1|state), data = sum1)
summary(m1)
ggpredict(m1, c("pct_goal","rounded")) |> plot()

m1 <- lmer(pct_change ~ refClass*pct_goal*rounded  + (1|id) + (1|state), data = sum1)
ggpredict(m1, c("refClass","pct_goal","rounded")) |> plot()

m1 <- lmer(pct_change ~ state*pct_goal*rounded  + (1|id), data = sum1)
summary(m1)
ggpredict(m1, c("state","pct_goal","rounded")) |> plot()

# m1 <- lmer(pct_change ~ refClass*pct_goal*rounded*state  + (1|id), data = sum1)
# ggpredict(m1, c("refClass","pct_goal","rounded","state")) |> plot()

@@

@@

# sum1 |> 
#   ungroup() |> 
#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
#   mutate(id=reorder(id,pct_change)) |> 
#   ggplot(aes(y=id,x=pct_change)) + 
#   geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
#   geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
#   ggh4x::facet_nested_wrap(refClass~pct_goal,axes="all",scales="free",ncol=2)
